---
title: "US. Doctoral Respondents Estimation through 2022 ACS"
author: 
  - Ariel Xing
  - Haoan Wang
thanks: "Code and data are available at: https://github.com/Jiaqi-Xing/US_Doctoral_Analysis."
date: today
date-format: long
abstract: ""
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(readr)
library(tidyverse)

clean <-read_csv("../data/analysis_data/cleaned_data.csv")
raw<-read_csv("../data/usa_00001.csv")
```

# Introduction 
Our data is obtained from [@ipums2022]. The dataset was cleaned and processed using R [@citeR], with help of packages tidyverse [@tidyverse], readr[@readr]. The cleaning process involved removing any unnecessary variables, and calculating the LaPlace estimations.

## Instructions for Extracting Data from IPUMS USA
1. Access the IPUMS USA Website
  Visit IPUMS USA.
2. Search for Variables
   In the search bar, enter the following variables one at a time:
   EDUC (Education)
   SEX (Gender)
   STATEICP (State, using the ICP coding system)
3. Add Variables to Your Cart
   After searching for each variable, select it and add it to your cart.
4. Create Data Extract
   Once all desired variables are in your cart, proceed to create the data extract.
5. Change Data Format
   In the extract request page, select CSV as the data format. This will make it easier for you to work with the data.
6. Submit Your Request
   Review your selections and submit the data extract request.
7. Download Your Data
   After your request is processed, navigate to the download page at Download Data. Locate your data file, which will be named usa_00001.csv, and download it.

# Overview of Ratio Estimator
First, we start with the how many respondents have a doctor degree as their highest educational attainment for each state. See "Table 1" for overview.
```{r}
#| label: dc_overview
#| fig-cap: 'Table 1: Sample of doctor degree for each state'
#| echo: false
# use 'knitr' generate a table of cleaned data for overview
knitr::kable(head(clean[, 1:2]), "simple",   col.names = c('ID', 'Counts'))

```


Then, we find the ratio for California, which has the ID 71 given by codebook. After that we calculate the ratio of doctoral degree respondents to total respondents in California by number of respondents with doctoral degrees in California divided by total number of respondents in California, which is given by the teacher(391,171). Then, use the ratio, we can calculate the number of respondents for every state by number of doctor degree divided by the ratio. See table 2 for overview of the estimated total respondents.

```{r}
#| label: dc_estimate
#| fig-cap: 'Table 2: Sample of doctor degree for each state'
#| echo: false
doctoral_degrees_california <- clean %>%
  filter(STATEICP == 71) %>%
  pull(respondents) 
total_respondents_california <- 391171
ratio_california <- doctoral_degrees_california / total_respondents_california
state_edu_data <- clean %>%
  mutate(estimated_total_respondents = respondents / ratio_california)
knitr::kable(head(state_edu_data[, 1:3]), "simple",   col.names = c('ID', 'Counts','Estimated'))
```

Then we compare with the actual number.
```{r}
#| label: dc_compare
#| fig-cap: 'Table 3: Sample of estimate v.s. actula'
#| echo: false
actual_respondents <- raw %>%
  group_by(STATEICP) %>%
  summarise(actual_total_respondents = n())
comparison_data <- state_edu_data %>%
  left_join(actual_respondents, by = "STATEICP") 

# Step 3: Calculate the difference between estimated and actual respondents
comparison_data <- comparison_data %>%
 select(STATEICP, respondents, estimated_total_respondents, actual_total_respondents)
knitr::kable(head(comparison_data[, 1:4]), "simple",   col.names = c('ID', 'Counts','Estimated','Actual'))
  
```

# Explaination of Difference
The Laplace ratio estimator is based on the ratio of doctoral degree holders in California to the total respondents in California, applied to other states. This method assumes that the proportion of doctoral degree holders is similar across states, which may not always be true. 
Each state has different demographics, education systems, and labor markets, leading to variations in educational attainment levels. For instance, states with a higher concentration of research institutions or universities may have more doctoral degree holders.
If certain respondents in a state did not report their education level, this could lead to undercounts in the actual totals.
The estimation assumes that the ratio of doctoral degree holders to total respondents is consistent across states. If certain states have unique educational trends or policies affecting the attainment of doctoral degrees, the estimate may not hold true.


\newpage


# References


